---
title: "Plots and statistics for: Putting in the effort: Predicting communicative work based on joint communication effort and conceptual affordances (preliminary results and pre-registration)"
output:
  html_document:
date: "2023-04-03"
---

```{r include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
```


```{r include=FALSE,warning=FALSE, message= FALSE}
library(readr)        #string manipulation
library(ggplot2)      #plotting
library(gridExtra)    #plotting
library(cowplot)      #plotting
library(colorRamps)   #plotting
library(RColorBrewer) #plotting
library(scales)       #rescaling
library(plotly)       #plotting
library(ggpubr)       #plotting
library(magrittr)     #pipe operator
library(lme4)         #lineat mixed modeling
library(ggcorrplot)   #plotting correlations
library(data.table)   #data wrangling
library(reshape2)     #data wrangling
library(plyr)         #data wrangling
library(dplR)         #data wrangling
library(dplyr)        #data wrangling
library(lmerTest)     #posthoc testing
library(simr)         #power analysis
library(tidyr)        #piping etc.
library(mixedpower)   #power analysis #devtools::install_github("DejanDraschkow/mixedpower")
library(performance)  #model checks
library(jtools)       #for plotting interaction
library(ape)          #ape
```

This document contains all the plots and statistical testing for the preregistration. Every header with a confirmatory label is considered the prime analyses to be replicated in the large confirmatory dataset. The other analyses labeled exploratory we will perform but should be treated as such.

## setup dir & read processed datasets
```{r, warning = FALSE, message = FALSE}
curfolder <- getwd()
outputfolder <- paste0(curfolder, "/processed/")

# loglog/linear LMs intercepts/slopes
final_mat<-read.csv(paste0(outputfolder,"mat_lm_prepost_centered_shift_one.csv")) 
final_mat_linear <- read.csv(paste0(outputfolder,"mat_lm_prepost_centered_linear.csv")) 

# raw sum(energy) per pair
energy <- read.csv(paste0(outputfolder,"energy.csv"))

# raw energy per participant
mat_diff <- read.csv(paste0(outputfolder,"mat_diff.csv"))

# pre/post/visual centrality, average all the rest Fribbles for each Fribble (result in 1 x 16)
centrality_all <- read.csv(paste0(outputfolder,"centrality_all.csv"))

# centrality in long format: post and pre distances
rdm_pair_long <- read.csv(paste0(outputfolder,"rdm_pair_long.csv"))

# visual similarity 16 x 16 Fribbles
vis_sim <- read.csv(paste0(outputfolder,"Ssim_RDM.csv"),header = F) 
vis_sim <- vis_sim[1:16,1:16]
vis_sim[row(vis_sim)==col(vis_sim)] <- NA

# post/pre centrality
rdm_mat <- read.csv(paste0(outputfolder,"rdm_mat_post.csv"))
rdm_mat_pre <- read.csv(paste0(outputfolder,"rdm_mat_pre.csv"))
```


## linear vs. loglog plot: for average efforts and 4 measures seperately

```{r, warning = FALSE, message = FALSE}
tmp <- mat_diff
tmp$File <- NULL

tmp<- tmp %>% group_by(round,pair) %>% summarise(avg_env=mean(peaks1+peaks2),avg_ges=mean(gesture1+gesture2),avg_trans=mean(transitions),avg_f0range=mean(range1+range2))

tmp[,c(3:6)] <- scale(tmp[,c(3:6)])

tmp$avg_effort <- rowMeans(tmp[,c(3:6)])

p1 <- ggplot(tmp,aes(x=round,y=avg_effort)) + geom_point(aes(colour=as.factor(pair), group=as.factor(pair)))+ guides(colour="none")+geom_smooth(method='lm', color='black', alpha=0.1)+geom_smooth(method='loess', color='purple', alpha=0.1)+ylab("combined effort")+theme_cowplot()

p2 <- ggplot(tmp,aes(x=log(round+1),y=log(avg_effort+1))) + geom_point(aes(colour=as.factor(pair), group=as.factor(pair)))+geom_smooth(method='loess', color='purple', alpha=0.1)+geom_smooth(method='lm', color='black', alpha=0.1)+ylab("log combined effort")+ theme(legend.position="right")+scale_colour_discrete(name="Pair")+theme_cowplot()

ggarrange(p1,p2,common.legend = TRUE,legend = "bottom")

```

# Explained variance log-log versus untransformed
```{r, warning = FALSE, message = FALSE}
library(MuMIn)
tmp <- mat_diff
tmp$File <- NULL

tmp<- tmp %>% group_by(round,pair) %>% summarise(avg_env=mean(peaks1+peaks2),avg_ges=mean(gesture1+gesture2),avg_trans=mean(transitions),avg_f0range=mean(range1+range2))

tmp[,c(3:6)] <- scale(tmp[,c(3:6)])

tmp$avg_effort <- rowMeans(tmp[,c(3:6)])

#test R squared for mixed regression model
modelunt <- lmer(avg_effort~round+(1|pair),data=tmp) 
summary(modelunt)

tmp$avg_effortl <- log(tmp$avg_effort+1)
tmp$roundl <- log(tmp$round)
#test R squared for mixed regression model
modellog <- lmer(avg_effortl~roundl+(1|pair),data=tmp) 
summary(modellog)

#Determine R2:
r.squaredGLMM(modelunt) 

#Determine R2:
r.squaredGLMM(modellog) 
```
## Plot untransformed and log log
```{r warning = FALSE, message = FALSE}
p1 <- ggplot(tmp,aes(x=scale(round),y=avg_ges)) + geom_point(aes(colour=as.factor(pair), group=as.factor(pair)))+ guides(colour="none")+geom_smooth(method='loess', color='black')+ylab("effort")

p2 <- ggplot(tmp,aes(x=scale(log(round)),y=log(avg_ges+1))) + geom_point(aes(colour=as.factor(pair), group=as.factor(pair)))+geom_smooth(method='lm', color='black')+ylab("log shift effort")+ theme(legend.position="right")+scale_colour_discrete(name="Pair")

p <- ggarrange(p1,p2,common.legend = TRUE,legend = "bottom")

annotate_figure(p,top=text_grob("gesture change over rounds"))

```
## correlations: centrality of Fribbles (HYPOTHESIS 1)
Visual similarity VS pre/post naming distance (per pair). Visdist correlate more with pre than with post naming distance (as expected)
```{r}
ggcorrplot(cor(centrality_all[,2:4]),method="square",lab=T)
```

### Mantel test: conclusion the same as avg per Fribble (the fig above)

Pre naming vs visual dissimilarity (main confirmatory: Hypothesis 1)

```{r}
mantel.test(rdm_mat_pre,vis_sim) # set graph = T
```

```{r, warning = FALSE, message = FALSE}
ggplot(centrality_all,aes(x=visdist,y=pair_predist)) + geom_point()+geom_smooth(method='lm', color='black')+ylab("pre centrality distance")+xlab("visual distance")
```

Post naming vs visual dissimilarity (exploratory)
```{r}
mantel.test(rdm_mat,vis_sim)
```

Pre naming vs post naming (exploratory)
```{r}
mantel.test(rdm_mat_pre,rdm_mat)
```

Pre  naming vs. post naming (exploratory)
```{r}
# double-check with vegan package - similar result
mantel.test(rdm_mat_pre,rdm_mat,na.rm = T) # 0.001 0.004 0.026
```
## Prepare data for work analyses
```{r, warning = FALSE, message = FALSE}
tmp_mat <- final_mat
tmp_mat[,c(2:9,11:14)] <- scale(tmp_mat[,c(2:9,11:14)])
tmp_mat$avg_intercepts <- rowMeans(tmp_mat[,c("trans_intercept","env_intercept","ges_intercept","f0_intercept")])
tmp_mat$avg_slopes <- rowMeans(tmp_mat[,c("trans_slope","env_slope","ges_slope","f0_slope")])
all <- left_join(tmp_mat[,c(1:16)],energy[,c(1,2,8)],by=c("pair","target"))
all<- left_join(all,rdm_pair_long,by=c("pair","target"="fribble"))
```

## plot: change/convergence ~ avg_slopes (confirmatory)

```{r, warning = FALSE, message = FALSE}
p1 <- ggplot(all,aes(x=avg_slopes,y=avg_dist)) + geom_point(aes(colour=factor(target), group=pair))+geom_smooth(method='lm', color='black')+ylab("change")+ggtitle("Average slope and change")

p1
```

# Power analyses slopes and change (Confirmatory hypothesis 2)
```{r, warning = FALSE, message = FALSE}
model1 <- lmer(prepost_change~(1|target)+(1|pair)+avg_slopes,data=all) 

summary(model1)


#data based power analyses
  #devtools::install_github("DejanDraschkow/mixedpower") 
  #if (!require("devtools")) {
  #   install.packages("devtools", dependencies = TRUE)}
  #   devtools::install_github("DejanDraschkow/mixedpower") 

power_model1 <- mixedpower(model = model1, data = all,
                      fixed_effects = c("avg_slopes"),
                      simvar = "pair", steps = c(42, 47),
                      critical_value = 1.96, n_sim = 500)
power_model1
```

## plot: intercepts x predist
```{r, warning = FALSE, message = FALSE}
res <- all
res$three <- cut(rank(res$predist),breaks=3,labels = c("low predist","mid predist","high predist")) 


p<-ggplot(res, aes(x=avg_intercepts,y=avg_dist)) + geom_point(aes(colour=factor(target), group=pair))+geom_smooth(method='lm', color='black')+ylab("change")+facet_grid(.~three)+ggtitle("Average intercept and change for different Fribble pre naming distances")

p # facet_grid(three.~) if horizontal

# ggsave("./figs/three.png",p,height=7,width=7)
```

# Modeling Power analyses intercepts x predist (Confirmatory hypothesis 3)
```{r, warning = FALSE, message = FALSE}
model2 <- lmer(prepost_change~avg_intercepts*predist+(1|pair)+(1|target),data=all) 

summary(model2)

#data based power analyses
  #devtools::install_github("DejanDraschkow/mixedpower") 
  #if (!require("devtools")) {
  #   install.packages("devtools", dependencies = TRUE)}
  #   devtools::install_github("DejanDraschkow/mixedpower") 

power_model2 <- mixedpower(model = model2, data = all,
                      fixed_effects = c("avg_intercepts", "predist"),
                      simvar = "pair", steps = c(42, 47),
                      critical_value = 1.96, n_sim = 500)
power_model2
```

